---
title: "Stat 508 - Final Project"
output:
  pdf_document: 
    toc: true
  html_notebook: default
  html_document:
    df_print: paged
author:
- Adam Behrendorff
- Nicholas Napier
- Saqib Ali
---


```{r}
suppressMessages(library(lubridate))
suppressMessages(library(caret))
suppressMessages(library(corrplot))
suppressMessages(library(sugrrants))
suppressMessages(library(dplyr))
suppressMessages(library(MASS))
suppressMessages(library(e1071))
suppressMessages(library(tree))


```

# Introduction

# Data

```{r}
EnergyDataAggregate <- read.csv("EnergyDataAggregate.csv",stringsAsFactors = FALSE)
EnergyDataAggregate$DATE <- as.Date(EnergyDataAggregate$DATE,format="%m/%d/%Y")
EnergyDataAggregate$Weekend <- if_else(EnergyDataAggregate$DAYNAME %in% c("Saturday", "Sunday"), "Weekend", "Weekday")

```


## Time-series Trend

```{r}
p <- EnergyDataAggregate %>%  
  frame_calendar(x = HOUR, y = GENERAL_SUPPLY_KWH, date = DATE) %>% 
  ggplot(aes(x = .HOUR, y = .GENERAL_SUPPLY_KWH, group = DATE, colour = Weekend)) +
  geom_line() +
  theme(legend.position = "bottom")
prettify(p)
```


# Analysis

## Binary Response Variable
Since we working on predicting a High Consumption vs. Low Consumption, we will make a HighEvergyUse binary variable based on the mean GENERAL_SUPPLY_KWH for the hour in the day

```{r}
EnergyDataAggregate$HighEnergyUse <- as.factor(ifelse(EnergyDataAggregate$GENERAL_SUPPLY_KWH>mean(EnergyDataAggregate$GENERAL_SUPPLY_KWH),1,0))
EnergyDataAggregate$MONTH <- as.factor(EnergyDataAggregate$MONTH)
EnergyDataAggregate$DAY <- as.factor(EnergyDataAggregate$DAY)

```


## Spliting the Dataset

We will split the dataset into training and testing splits. We will use all the data from before 2013 to build a model to predict Evergy Consumption for years 2013 and higher.

```{r}
set.seed(1)

#training and test set
energyData.full <- EnergyDataAggregate
energyData.train=EnergyDataAggregate[EnergyDataAggregate$YEAR<2013,]
energyData.test=EnergyDataAggregate[EnergyDataAggregate$YEAR>2013,]

```




## Logistic Regression

Let's build a Logistic Model using the training Data
```{r}

logit.fit <- glm(HighEnergyUse~Weekend+DAY+MONTH+DAYNAME+HOUR, data=energyData.train, family=binomial)


```



```{r}
glm.probs=predict(logit.fit,energyData.test,type="response")
glm.pred <- ifelse(glm.probs<0.5, 0, 1)

pred.mean <- mean(glm.pred==energyData.test$HighEnergyUse)
pred.mean
```
We observe that with Logistic Regression, we get and accurary rate of `r pred.mean`

## LDA
```{r}
lda.fit=lda(HighEnergyUse~Weekend+DAY+MONTH+DAYNAME+HOUR, data=energyData.train)
```


```{r}
lda.pred=predict(lda.fit, energyData.test)
pred.mean <- mean(lda.pred$class==energyData.test$HighEnergyUse)
```

We observe that with LDA, we get and accurary rate of `r pred.mean`


## QDA

### With Day of the Month

```{r}
qda.fit=qda(HighEnergyUse~DAY+MONTH+DAYNAME+HOUR, data=energyData.train)
```


```{r}
qda.pred=predict(qda.fit, energyData.test)
pred.mean <- mean(qda.pred$class==energyData.test$HighEnergyUse)
```
We observe that with QDA, we get and accurary rate of `r pred.mean`


### Without Day of the Month
```{r}
qda.fit=qda(HighEnergyUse~DAY+MONTH+DAYNAME+HOUR, data=energyData.train)
```


```{r}
qda.pred=predict(qda.fit, energyData.test)
pred.mean <- mean(qda.pred$class==energyData.test$HighEnergyUse)
```
We observe that with QDA, we get and accurary rate of `r pred.mean`




## Support Vector Classifer
```{r}
tune.cost <- c(0.001, 0.01, 0.1, 1,5,10,100)

tune.out=tune(svm, HighEnergyUse~DAY+MONTH+DAYNAME+HOUR,data=energyData.full,kernel="linear",ranges=list(cost=tune.cost))


```

```{r}
bestmod=tune.out$best.mode
ypred=predict(bestmod ,energyData.train)
mean(ypred==energyData.train$HighEnergyUse)
```



## Regression Trees

```{r}
EnergyData.tree <- tree(HighEnergyUse~READING_EPOCHTIME+Weekend+DAY+MONTH+DAYNAME+HOUR, energyData.train)
 plot(EnergyData.tree) 
 text(EnergyData.tree ,pretty=0)
```

```{r}
 set.seed(3) 
tree.pred=predict(EnergyData.tree,energyData.test,type="class") 
table(tree.pred, energyData.test$HighEnergyUse)
mean(tree.pred==energyData.test$HighEnergyUse)

```



